---
title: "Exercisis 3. Inferencia"
author: "Magí Bas"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Faraway

1.  (Ejercicio 1 cap. 3 pág. 48) For the prostate data, fit a model with
    lpsa as the response and the other variables as predictors:

    \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

    **Description** The prostate data frame has 97 rows and 9 columns. A
    study on 97 men with prostate cancer who were due to receive a
    radical prostatectomy.

    **Format** This data frame contains the following columns:

    -   lcavol log(cancer volume)

    -   lweight log(prostate weight)

    -   age age

    -   lbph log(benign prostatic hyperplasia amount)

    -   svi seminal vesicle invasion

    -   lcp log(capsular penetration)

    -   gleason Gleason score

    -   pgg45 percentage Gleason scores 4 or 5

    -   lpsa log(prostate specific antigen)

    \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

```{r}
library(faraway)
library(tidyverse)
#library(tidyverse)
data(prostate)

m=lm(lpsa~., data=prostate)
summary(m)

```

(a) **Compute 90 and 95% CIs for the parameter associated with age.
    Using just these intervals, what could we have deduced about the
    p-value for age in the regression summary?**

```{r}
#Intervalos de confianza
confint(m, c("age"), .95)
confint(m, c("age"), .90)
```

Con base en los intervalos de confianza 'age' está en el límite para ser
considerado significativo. Debido a que el IC del 90 % no incluye el 0,
la variabe "age" es significativa en ese nivel, pero el límite superior
del IC del 95 %, que es más estricto, se extiende justo por encima de
cero hasta 0,0025.

En resumen de regresión lo confirma al devolver un valor p de 0,08,
cercano pero superior al umbral común de 0,05 para la significación.

**b. Compute and display a 95% joint confidence region for the
parameters associated with age and lbph. Plot the origin on this
display. The location of the origin on the display tells us the outcome
of a certain hypothesis test. State that test and its outcome**

```{r prostate_b}
library(ellipse)
plot(ellipse(m, c('age', 'lbph')), type = "l")
points(0, 0, pch = 1)
abline(v= confint(m)['age',], lty = 2)
abline(h= confint(m)['lbph',], lty = 2)
```

La hipótesis nula conjunta \`age = lbph = 0\` no se puede rechazar
porque el origen se encuentra dentro de la elipse de la región de
confianza. De manera similar, la hipótesis nula \`age = 0\` no se puede
rechazar porque 0 se encuentra dentro de los límites de confianza del
95% y lo mismo es cierto para la hipótesis nula \`lbph = 0\`.

**(c) In the text, we made a permutation test corresponding to the F
-test for the significance of all the predictors. Execute the
permutation test corresponding to the t-test for age in this model.
(Hint: summary(g)\$coef[4,3] gets you the t-statistic you need if the
model is called g.)**

```{r}
library(magrittr)
library(purrr)
library(mlbench)
t_value = summary(m) %>% coef() %>% .['age', 't value']

#funcion para permutar n-veces
permute_tmod = function(nsims) {
  map_dbl(1:nsims,
          ~ lm(sample(lpsa) ~., data=prostate) %>%
            summary() %>%
            coef() %>%
            .['age', 't value'])
}

mean(abs(permute_tmod(100)) > abs(t_value))
mean(abs(permute_tmod(1000)) > abs(t_value))
mean(abs(permute_tmod(10000)) > abs(t_value))
```

De la sección a anterior, no sabemos que el valor p para \`edad\` es
0.08229, y podemos ver que el valor de retorno de la permutación se
acerca a ese número a medida que aumenta el número de simulaciones.

**(d) Remove all the predictors that are not significant at the 5%
level. Test this model against the original model. Which model is
preferred?**

```{r}
m2 <- update(m, . ~ lcavol + lweight + svi)
anova(m, m2)
```

El modelo reducido no es significativamente mejor que el modelo
completo, por lo que elegiríamos \`m\` sobre \`m2\`.

**2. (Ejercicio 2 cap. 3 pág. 49) Thirty samples of cheddar cheese were
analyzed for their content of acetic acid, hydrogen sulfide and lactic
acid. Each sample was tasted and scored by a panel of judges and the
average taste score produced. Use the cheddar data to answer the
following:**

```{r}
data(cheddar)
str(cheddar)
```

**(a) Fit a regression model with taste as the response and the three
chemical contents as predictors. Identify the predictors that are
statistically significant at the 5% level.**

```{r}
cheddar_model=lm(taste~., data=cheddar)
summary(cheddar_model)
```

The predictors that are satististically significant are: hydrogen
sulfide and lactic acid

(b) Acetic and H2S are measured on a log scale. Fit a linear model where
    all three predictors are measured on their original scale. Identify
    the predictors that are statistically significant at the 5% level
    for this model.

```{r}
delog_cheddar_model=update(cheddar_model, . ~exp(Acetic)+exp(H2S)+Lactic)
summary(delog_cheddar_model)
```

Now, only Lactic is statistically significant a 5%

**(c) Can we use an F -test to compare these two models? Explain. Which
model provides a better fit to the data? Explain your reasoning.**

```{r}
# F-test are used to acess variation (expected/un-expected) in a model
anova(cheddar_model, delog_cheddar_model)
```

There is less residual variance in the log model (m), but we don't have
any degrees of freedom left so we can't estimate the F-statistic

**(d) If H2S is increased 0.01 for the model used in (a), what change in
the taste would be expected?**

```{r}
(coef(cheddar_model)[3]*.01)
```

0.039 increase in taste

**(e) What is the percentage change in H2S on the original scale
corresponding to an additive increase of 0.01 on the (natural) log
scale?**

```{r}
test_df = data.frame(Acetic = rep(mean(cheddar$Acetic), 2),
                     H2S=mean(cheddar$H2S)+c(0,.01),
                     Lactic = mean(cheddar$Lactic))
predict(cheddar_model, test_df)
```

confirmed

```{r}
(coef(cheddar_model)[3] + 0.01)/coef(delog_cheddar_model)[3]
```

5114% increase to match

```{r}
exp(.01)
```

1 point increse in the unlogged scale

**3. (Ejercicio 3 cap. 3 pág. 49) Using the teengamb data, fit a model
with gamble as the response and the other variables as predictors.**

```{r}
library(skimr)
?teengamb
skim(teengamb)
```

**(a) Which variables are statistically significant at the 5% level?**

```{r}
teengamb_model=lm(gamble ~., data=teengamb)
summary(teengamb_model)
```

Sex and income significant at 5% level

**(b) What interpretation should be given to the coefficient for sex?**

Since th coefficient shows a -22.11 shift in gamble for an increase of 1
in sex and females are code as 1, women gamble less than men

**(c) Fit a model with just income as a predictor and use an F -test to
compare it to the full model.**

```{r}
# if you recode sex as a factor with the levels as c(female, male)
teengamb$sex %<>% fct_rev()
# and refit
summary(teengamb_model ~ ., data=teengamb)
```

```{r}
summary(teengamb_model)
```

```{r}
teengamb_model2= lm(gamble ~ income, data=teengamb)
summary(teengamb_model2)
```

```{r}
# order of anova doesn't matter
#larger model is significanly better
anova(teengamb_model, teengamb_model2)
```
